A Comparison of Algorithms Playing EvoMan

### 2020-06-15 paper/paper.tex
sync 2.
Added final table.

    \subsection{Best Combination of Opponents for Training}\label{subsec:best-combination-of-opponents-for-training}
    We searched for the opponents that lead to the best results.
    The first combination of train opponents we considered was \{1, 2, 6, 7\} which was chosen empirically after manually playing against every opponent and chosing the ones which behaved in such a way that covered the most general concepts our agents could learn in our opinion.
    We observed its final gain as 125.37.
    We realized a set of exploratory experiments where we would search for other combinations of train opponents,
    but starting with the resulting model of PPO trained for 1000 epochs on the combination of opponents \{1, 2, 6, 7\}.
    If by starting with a pre-trained model another combination of train enemies does not lead to better results,
    then we can believe that the respective combination does not lead to better results than \{1, 2, 6, 7\}.
    \begin{table}[htbp]
        \caption{Results for various opponents when starting with a pre-trained PPO model on enemies \{1, 2, 6, 7\}}
        \begin{center}
            \begin{tabular}{|c|c|}
                \hline
                \textbf{Train opponents}&{\textbf{gain (harmonic mean)}} \\
                \cline{2-4}
                \hline
                \{1, 2, 6, 7\} \cup \{3, 4, 5, 8\} & 44.16 \\
                \{1, 2, 6, 7\} \cup \{3, 4, 6, 7\} & 75.3 \\
                \{1, 2, 6, 7\} \cup \{1, 3, 6, 7\} & 110.19 \\
                \{1, 2, 6, 7\} \cup \{2, 3, 6, 7\} & 136.1 \\
                \{1, 2, 6, 7\} \cup \{2, 4, 6, 7\} & 85.22 \\
                \hline
            \end{tabular}
            \label{table:variations_of_opponents}
        \end{center}
    \end{table}
    \\
    The only combination that lead to a better score than 125.37 was \{2, 3, 6, 7\}.
    We ran an experiment with PPO with random initialization with the combination of train enemies \{2, 3, 6, 7\} and
    we obtained a final gain of 99.97.
    This means that we did not find a better combination of train opponents than \{1, 2, 6, 7\}, so it is the one
    we used for the following experiments.

@@

@@

    \begin{thebibliography}{00}
        \bibitem{evoman_competition} Evoman: Game-playing Competition for WCCI 2020, \url{http://pesquisa.ufabc.edu.br/hal/Evoman.html}
        \bibitem{evoman} Fabricio Olivetti de Franca, Denis Fantinato, Karine Miras, A.E. Eiben and Patricia A. Vargas.
        "EvoMan: Game-playing Competition" arXiv:1912.10445
        \bibitem{karinemiras} de Araújo, Karine da Silva Miras, and Fabrício Olivetti de França.
        "An electronic-game framework for evaluating coevolutionary algorithms." arXiv:1604.00644 (2016).
        \bibitem{neuro} Floreano, D., Dürr, P. & Mattiussi, C. Neuroevolution: from architectures to learning. Evol. Intel. 1, 47–62 (2008). \url{https://doi.org/10.1007/s12065-007-0002-4}
        \bibitem{capcom} M. MEGA, "Produced by capcom, distributed by capcom, 1987," System: NES.
        \bibitem{q_learning} Watkins, C.J.C.H., Dayan, P. Q-learning. Machine Learning 8, 279–292 (1992)
        \bibitem{genetic_algorithm} Holland J.H., Genetic Algorithms and Adaptation. Adaptive Control of Ill-Defined Systems, 1984, Volume 16 ISBN 978-1-4684-8943-9
        \bibitem{pso} Kennedy, J.; Eberhart, R. (1995). "Particle Swarm Optimization". Proceedings of IEEE International Conference on Neural Networks. IV. pp. 1942–1948.
        \bibitem{neat} Kenneth O. Stanly; Rist Miikkulainen (2002). "Evolving Neural Networks through Augmenting Topologies". Evolutionary Computation, Volume 10, Issue 2, Summer 2002, p.99-127, \url{https://doi.org/10.1162/106365602320169811}
        \bibitem{ppo} John Schulman, Filip Wolski, Prafulla Dhariwal, Alex Radford, Oleg Klimov (2017) "Proximal Policy Optimization Algorithms", arXiv:1707.06347v2
        \bibitem{evoman_blog} Karine Miras, Evoman, \url{https://karinemirasblog.wordpress.com/portfolio/evoman/}
    \end{thebibliography}

>>>>

    \subsection{Best Combination of Opponents for Training}\label{subsec:best-combination-of-opponents-for-training}
    We searched for the opponents that lead to the best results.
    The first combination of train opponents we considered was \{1, 2, 6, 7\} which was chosen empirically after manually playing against every opponent and chosing the ones which behaved in such a way that covered the most general concepts our agents could learn in our opinion.
    We observed its final gain as 125.37.
    We realized a set of exploratory experiments where we would search for other combinations of train opponents,
    but starting with the resulting model of PPO trained for 1000 epochs on the combination of opponents \{1, 2, 6, 7\}.
    If by starting with a pre-trained model another combination of train enemies does not lead to better results,
    then we can believe that the respective combination does not lead to better results than \{1, 2, 6, 7\}.
    \begin{table}[htbp]
        \caption{Results for various opponents when starting with a pre-trained PPO model on enemies \{1, 2, 6, 7\}}
        \begin{center}
            \begin{tabular}{|c|c|}
                \hline
                \textbf{Train opponents}&{\textbf{gain (harmonic mean)}} \\
                \hline
                $\{1, 2, 6, 7\} \cup \{3, 4, 5, 8\}$ & $44.16$ \\
                $\{1, 2, 6, 7\} \cup \{3, 4, 6, 7\}$ & $75.3$ \\
                $\{1, 2, 6, 7\} \cup \{1, 3, 6, 7\}$ & $110.19$ \\
                $\{1, 2, 6, 7\} \cup \{2, 3, 6, 7\}$ & $136.1$ \\
                $\{1, 2, 6, 7\} \cup \{2, 4, 6, 7\}$ & $85.22$ \\
                \hline
            \end{tabular}
            \label{table:variations_of_opponents}
        \end{center}
    \end{table}
    \\
    The only combination that lead to a better score than 125.37 was \{2, 3, 6, 7\}.
    We ran an experiment with PPO with random initialization with the combination of train enemies \{2, 3, 6, 7\} and
    we obtained a final gain of 99.97.
    This means that we did not find a better combination of train opponents than \{1, 2, 6, 7\}, so it is the one
    we used for the following experiments.

@@

    \begin{table}[htbp]
        \caption{Average player and enemy life. Average game duration}
        \begin{center}
            \begin{tabular}{|c|c|c|c|c|c|c|}
                \hline
                \textbf{Opponent}&\multicolumn{3}{|c|}{\textbf{Best train agent}}&\multicolumn{3}{|c|}{\textbf{Best test agent}} \\
                \cline{2-7}
                & \textbf{\textit{Avg player life}}& \textbf{\textit{Avg enemy life}}& \textbf{\textit{Avg duration}} & \textbf{\textit{Avg player life}}& \textbf{\textit{Avg enemy life}}& \textbf{\textit{Avg duration}} \\
                \hline
                1 & 97.26 &   100 &     0 &     0 & 159.63 & 174.23 \\
                2 & 98.33 & 89.73 &     0 &     0 &  176.6 & 186.43 \\
                3 &  0.13 &  7.66 & 43.33 & 21.33 & 457.26 & 410.53 \\
                4 &     0 &  1.84 & 37.66 & 21.66 &  547.5 & 736.83 \\
                5 & 71.32 & 60.52 &     0 &     0 & 379.13 & 367.93 \\
                6 & 96.32 &    95 &     0 &     0 & 165.06 & 203.53 \\
                7 & 99.74 &  85.9 &     0 &     0 & 162.06 & 135.13 \\
                8 &  34.3 & 38.76 &     6 &     2 & 520.93 & 443.6 \\
                \hline
            \end{tabular}
        \end{center}
    \end{table}

@@

    \begin{thebibliography}{00}
        \bibitem{evoman_competition} Evoman: Game-playing Competition for WCCI 2020, \url{http://pesquisa.ufabc.edu.br/hal/Evoman.html}
        \bibitem{evoman} Fabricio Olivetti de Franca, Denis Fantinato, Karine Miras, A.E. Eiben and Patricia A. Vargas.
        "EvoMan: Game-playing Competition" arXiv:1912.10445
        \bibitem{karinemiras} de Araújo, Karine da Silva Miras, and Fabrício Olivetti de França.
        "An electronic-game framework for evaluating coevolutionary algorithms." arXiv:1604.00644 (2016).
        \bibitem{neuro} Floreano, D., Dürr, P. \& Mattiussi, C. Neuroevolution: from architectures to learning. Evol. Intel. 1, 47–62 (2008). \url{https://doi.org/10.1007/s12065-007-0002-4}
        \bibitem{capcom} M. MEGA, "Produced by capcom, distributed by capcom, 1987," System: NES.
        \bibitem{q_learning} Watkins, C.J.C.H., Dayan, P. Q-learning. Machine Learning 8, 279–292 (1992)
        \bibitem{genetic_algorithm} Holland J.H., Genetic Algorithms and Adaptation. Adaptive Control of Ill-Defined Systems, 1984, Volume 16 ISBN 978-1-4684-8943-9
        \bibitem{pso} Kennedy, J.; Eberhart, R. (1995). "Particle Swarm Optimization". Proceedings of IEEE International Conference on Neural Networks. IV. pp. 1942–1948.
        \bibitem{neat} Kenneth O. Stanly; Rist Miikkulainen (2002). "Evolving Neural Networks through Augmenting Topologies". Evolutionary Computation, Volume 10, Issue 2, Summer 2002, p.99-127, \url{https://doi.org/10.1162/106365602320169811}
        \bibitem{ppo} John Schulman, Filip Wolski, Prafulla Dhariwal, Alex Radford, Oleg Klimov (2017) "Proximal Policy Optimization Algorithms", arXiv:1707.06347v2
        \bibitem{evoman_blog} Karine Miras, Evoman, \url{https://karinemirasblog.wordpress.com/portfolio/evoman/}
    \end{thebibliography}
